---
Full name: Khang-Trang Nguyen
title: "Kritik-Updated-Assignment"
format:
  pdf:
    pdf-engine: xelatex
---

### The data measure the air quality every day through May to September

```{r, warning=FALSE, message=FALSE}
dat <- read.table("Student_Dataset_Assignment.txt", header = TRUE, sep = "\t")
data (airquality)
```

### Load and quick examine the data

```{r, warning=FALSE, message=FALSE}
dat <- airquality
str(dat)
summary(dat)
```

### So, there are some NAs in columns Ozone and Solar. R

### Check missing values carefully

```{r, warning=FALSE, message=FALSE}
colSums(is.na(dat))
```

### So, there are 37 NAs in column Ozone and 7 NAs in column Solar.R

### Examine Variables Wind vs Temp

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
#graph data
qplot(airquality$Wind, airquality$Temp)
cor(airquality$Wind, airquality$Temp)
qplot(airquality$Wind, airquality$Temp) + geom_smooth(method = "lm", se = FALSE)
model <- lm(Temp ~ Wind, data = airquality)
summary(model)
```

```{r, warning=FALSE, message=FALSE}
coef(model)
```

### For every unit that wind increases, the temperature will decrease by about 1.23 units.

## Residuals vs Fitted Values

```{r, warning=FALSE, message=FALSE}
qplot(fitted.values(model), residuals(model))
qplot(fitted.values(model), residuals(model))+geom_smooth(method = "lm", se = FALSE)
```

### There's no pattern to this data, all the residuals scatter around the 0 line; therefore, Wind vs Temp have a weak linear relationship (since the correlation coefficient is -0.46.

## Graph of correlation between Ozone level (y) vs Months (x)

```{r, warning=FALSE, message=FALSE}
library(ggplot2)

ggplot(airquality, aes(x = factor(Month), y = Ozone)) +
  geom_boxplot(na.rm = TRUE, fill = "lightblue") +
  labs(title = "Ozone levels by Month",
       x = "Month",
       y = "Ozone (ppb)") +
  theme_minimal()
```

### Conclusion: Ozone levels reach their highest values in August, while the lowest values occur in May and July.

## Graph of correlation between Solar. R (y) vs Months (x)

```{r, warning=FALSE, message=FALSE}
library(ggplot2)

ggplot(airquality, aes(x = factor(Month), y = Solar.R)) +
  geom_boxplot(na.rm = TRUE, fill = "orange") +
  labs(title = "Solar Radiation by Month",
       x = "Month",
       y = "Solar.R") +
  theme_minimal()
```

### Conclusion: Solar radiation is highest in June and July, and lowest in May and September.

## Graph of correlation between Wind (y) vs Months (x)

```{r, warning=FALSE, message=FALSE}
library(ggplot2)

ggplot(airquality, aes(x = factor(Month), y = Wind)) +
  geom_boxplot(na.rm = TRUE, fill = "lightgreen") +
  labs(title = "Wind Speed by Month",
       x = "Month",
       y = "Wind (mph)") +
  theme_minimal()
```

### Conclusion: Wind is strongest in May and weakest in August, with June showing moderate levels.

## Graph of correlation between Temp (y) vs Months (x)

```{r, warning=FALSE, message=FALSE}
library(ggplot2)

ggplot(airquality, aes(x = factor(Month), y = Temp)) +
  geom_boxplot(na.rm = TRUE, fill = "tomato") +
  labs(title = "Temperature by Month",
       x = "Month",
       y = "Temperature (°F)") +
  theme_minimal()
```

### Conclusion: Temperature is lowest in May but peaks in July and August.
